---
title: "Avocado Project"
author: "Evelyn Chen, Christen Paryzch"
date: "10/12/2019"
output: html_document
---
#Research topic: ‘Guac’-Bottom Exploring Avocados’ Fluctuating Prices

Loading libraries
```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(plotly)
library(Metrics)
library(forcats)
library(RColorBrewer)
library(forcats)
library(ModelMetrics)
library(modelr)
```

##Introduction
  The avocado is America’s ‘it’ fruit, so much so that it has become a running joke—headlines tell us that millennials choose avocados over, say, homeownership (Cummings, 2019). In spite of the hyperbole, the avocado has become an obsession Americans love to love. The evidence backs this claim: within the last week alone (October 6, 2019), U.S. consumers purchased 48,778,842 pounds of avocado—that’s a lot of avocado toast (Hass Avocado Board, 2019)! Mexico leads the world in avocado production, followed by the United States, where over 90% of avocado production occurs in California (Dekevich, 2018). 
  This exploratory report utilizes data from the Hass Avocado Board website, published in May 2018 and compiled into a CSV. The Hass Avocado Board is an agricultural advocacy group founded to promote the consumption of avocados within the United States. The dataset includes weekly retail scan data for national retail (grocery, mass, club, drug, dollar, and military) volume price, including 18,249 observations of 16 variables from 2015–2018. 
	We seek to explore a central question: are the price fluctuations in U.S. avocado-producing regions steadier/smaller than in non-avocado-producing regions? We are also interested in investigating whether the average price of Avocados is lower in avocado-producing regions. Finally, to what extent do region and avocado type predict the average price of avocados?

```{r}
rm(list=ls())
avocado<-read_csv(file="avocado.csv")
```

## Data
The data represents weekly retail scan data for National review volume and price, from 2015 - 2018. The retail scan data comes directly from retailers' cash registers, which reflects actual retail sales of Hass avocados. This is a quick explanation of the variables in the dataset.

Date: Date of observation
AveragePrice: the average price of a single avocado
Total Volume: total volume of avocados sold (pounds)
Total bags: total number of bags of avocados sold
Small bags: total number of small bags of avocados sold
Large bags: total number of large bags of avocados sold
XLarge Bags: total number of extra large bags of avocados sold
Type: conventional or organic
Year: the year
Region: the city or region of the sale
Total volume: total number of avocados sold
4046: total number of small/medium PLU 4046 avocados sold
4225: total number of large PLU 4225 avocados sold
4770: total number of extra-large PLU 4770 avocados sold

The data is largely consistent, with the exception of ‘region,’ which is a mix of cities, states, and regions of the country. We have elected to use regions of the country as our consistent characterization; we will need to clean up the data in line with the Hass Avocado Board’s existing regional categorizations for the continental U.S.:

•	California
•	West – Washington, Oregon, Idaho, Nevada, New Mexico, Montana, Colorado, Utah, Arizona, Wyoming
•	South Central – Texas, Oklahoma, Arkansas, Louisiana
•	Plains – North Dakota, South Dakota, Nebraska, Minnesota, Iowa, Kansas, Missouri
•	Great Lakes – Wisconsin, Indiana, Illinois, Ohio, Michigan
•	Southeast – Alabama, Mississippi, Georgia, South Carolina, Florida
•	Mid-South – Tennessee, Kentucky, North Carolina, Virginia, West Virginia, Maryland
•	Northeast – Maine, Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York, Pennsylvania, New Jersey, Delaware

Because we will be exploring price fluctuations in various regions of the country, we expect our analysis to focus on the following variables: average price, size of avocado, type, year purchased, and region purchased. Based on the available data, we will first identify a normal range for avocado prices according to region and then determine when avocado prices fell outside of this range from 2015 to 2018. The Hass Avocado Board provides rich and accessible data, so depending on the final scope of our project, we might integrate more in-depth, region-specific analyses into our report as well. 


### Tidying Data
```{r}
#Rename Columns
colnames(avocado)[colnames(avocado)=="Total Volume"]<- "TotalVolume"
colnames(avocado)[colnames(avocado)=="Total Bags"]<- "TotalBags"
colnames(avocado)[colnames(avocado)=="Small Bags"]<- "SmallBags"
colnames(avocado)[colnames(avocado)=="Large Bags"]<- "LargeBags"
colnames(avocado)[colnames(avocado)=="XLarge Bags"]<- "XLargeBags"
```


```{r}


avocado%>%rename(region="Albany")%>%mutate(region="Northeast")

```

##Exploratory Data Analysis

Univariate graphics helps to understand how individual variables are distributed across the sample. 
Below shows the frequency plot of avocado type, showing that this dataset includes retail sales of both conventional and organic avocados. We can use geom_bar to get this done. By default, this will give us a count in each type of avocado. 

```{r}
gg<-ggplot(avocado,aes(x=type,fill=type))
gg<-gg+geom_bar()
gg
```

The dataset shows weekly retail data of avocado sales from 2015 - 2018. We are also interested in the distribution of the average price of avocados in this dataset. We will use the ggplot command to create a density plot of the 'AveragePrice' variable. 
```{r}
gg<-ggplot(avocado,aes(x=AveragePrice,y=TotalVolume),color=as.factor(year))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+xlab("Average Price")+ylab("Total Volume")
gg
```



```{r}
avocado%>%summarize(mean_AveragePrice=mean(AveragePrice,na.rm=TRUE))
##unconditional average as a new variable
avocado<-avocado%>%mutate(mean_AveragePrice=mean(AveragePrice,na.rm=TRUE))
```

The distribution is fairly normal. The unconditional mean of 'AveragePrice' is 1.41. 

```{r error_terms}
avocado<-avocado%>%mutate(e1=AveragePrice-mean_AveragePrice)
rmse_averageprice<-rmse(avocado$AveragePrice,avocado$mean_AveragePrice)
rmse_averageprice
```
On average, we would be off by $0.40 across the whole dataset. At first glance, this error may not seem very big, but 'AveragePrice' ranges from 0.44 to 3.25. A RMSE error value of $0.40 is fairly significant, in this context. 




